## Mapping Vulnerability

### The final dataset

The final vulnerability scores were derived from `names(coins["Agg-gmean"])`, as this approach was considered to better represent species risk within the pilot areas. These species-level vulnerability scores were integrated with species distribution model (SDM) predictions and combined into a multi-layer *SpatRaster* object for subsequent spatial analyses.

```{r thefinalcoin}
finalcoin <- coins[["Agg-gmean"]]

tar_source(fs::path(here::here(), "..", "envBird", "R"))

final_sp <- finalcoin$Data$Aggregated %>% 
  dplyr::select(uCode, Vulnerability) %>% 
  dplyr::mutate(uCode = gsub("_", " ", uCode)) %>% 
  replace_taxa(taxa_col = "uCode", direction = "reverse")
```

### Vulnerability map - max method

The maximum value across the raster stack was extracted for each pixel to represent the highest species-level risk at that location. The resulting map therefore reflects the maximum vulnerability score observed per pixel.

```{r vulndf}
## paths joining vulnerability
usg_sp <- usg_sp %>% dplyr::inner_join(final_sp, by = c("taxa" = "uCode"))
bp_sp <- bp_sp %>% dplyr::inner_join(final_sp, by = c("taxa" = "uCode"))
```

```{r calculateraster}

dir = "/projects/dev/out/RiskMap/rasters"

# Define inputs
rast_inputs <- list(
  usg = list(rast_stack = usg_rasts, df = usg_sp),
  bp  = list(rast_stack = bp_rasts,  df = bp_sp)
)

agg_funs <- c("max", "sum")

# Compute all combinations
sp_rasts <- purrr::map(agg_funs, function(fun) {
  purrr::map(rast_inputs, function(inp) {
    calculate_vulnerability_surface(
      rast_stack = inp$rast_stack,
      df         = inp$df,
      agg_fun    = fun
    )
  })
}) %>%  
  purrr::set_names(agg_funs)

```

**The vulnerability map of Upper Spencer Gulf - Gawler Ranges area**

```{r usgmap, fig.width=10, out.width="100%"}
terra::plot(sp_rasts$max$usg,
            plg = list(title= "Vulnerability\nscore"))
```

**The vulnerability map of Braemar Province area**

```{r bpmap, fig.width=10, out.width="100%"}
terra::plot(sp_rasts$max$bp,
            plg = list(title= "Vulnerability\nscore"))
```

#### Apply zoning thresholds {-}

Theoretically, all species of conservation concern should fall within either the investigation (yellow) or no-go (red) zones. Therefore, under the current scenario where non-concerned species are not included, a threshold of 0 was applied to visualise the extent of the good-to-go (green) zone. The no-go (red) zone was defined using a threshold of 14, based on species with vulnerability scores near this boundary, including the Slender-billed Thornbill (14.1), Beautiful Firetail (13.5), and Chestnut-rumped Heathwren (15.0).

```{r maxzonemap}
purrr::walk2(
  sp_rasts$max,
  names(sp_rasts$max),
  function(rast, nm) {
    terra::plot(
      rast,
      breaks = c(0, 0.1, 14, Inf),
      col    = c("forestgreen", "yellow", "red"),
      main = paste0(toupper(nm), " — max: traffic light"),
      plg  = list(title = "Vulnerability\nscore")
    )
  }
)
```

### Vulnerability map - sum method

The sum method aggregates the vulnerability scores of all species predicted to occur within a location and highlights the hot spot where multiple species of conservation concern.

```{r sumzonemap}

purrr::walk2(
  sp_rasts$sum,
  names(sp_rasts$sum),
  function(rast, nm) {
    terra::plot(
      rast,
      main = paste0(toupper(nm), " — sum"),
      plg  = list(title = "Vulnerability\nscore")
    )
  }
)
```

### Three-Tier Zoning

Key considerations for applying a traffic-light (three-tier) zoning approach include:

1. All species, including those not currently of concern, need to be assessed to derive meaningful thresholds; this is particularly challenging for data-deficient groups.

2. Even with complete assessments, threshold selection remains inherently arbitrary, although it can be adjusted to reflect management priorities (e.g. a precautionary, risk-averse approach).

3. Allowing 0 to serve as the first threshold, representing the absence of species of conservation concern, simplifies the zoning framework by requiring only a single additional threshold to define the no-go (red) zone.

4. Using the maximum species-level risk to represent each pixel may oversimplify patterns, as it obscures variation among species and underlying biodiversity values.

```{r scoredistribution, fig.cap="Distribution of final vulnerability scores and the application of threshold breaks for three-tier zoning."}

## Natural breaks
breaks <- classInt::classIntervals(
  final_sp$Vulnerability, 
  n = 2, 
  style = "fisher"
)

# Extract just the two threshold values (drop min and max)
threshold <- breaks$brks[c(-1, -length(breaks$brks))]

# Count species per category
cat_counts <- final_sp %>%
  dplyr::mutate(
    VulnerabilityClass = dplyr::case_when(
      Vulnerability <= threshold[1] ~ "Moderate",
      TRUE                          ~ "High"
    )
  ) %>%
  dplyr::count(VulnerabilityClass) %>%
  # Order by vulnerability level, not alphabetically
  dplyr::mutate(
    VulnerabilityClass = factor(VulnerabilityClass, 
                                levels = c("Moderate", "High"))
  ) %>%
  dplyr::arrange(VulnerabilityClass) %>%
  dplyr::mutate(
    x = c(
      mean(c(min(final_sp$Vulnerability), threshold)),
      mean(c(threshold, max(final_sp$Vulnerability)))
    )
  )

ggplot2::ggplot(final_sp, ggplot2::aes(x = Vulnerability)) +
  ggplot2::geom_histogram(bins = 20, fill = "grey") +
  ggplot2::geom_density(ggplot2::aes(y = ggplot2::after_stat(count)), 
                        colour = "red") +
  ggplot2::geom_vline(
    xintercept = threshold,
    colour     = "steelblue",
    linetype   = "dashed",
    linewidth  = 0.8
  ) +
  ggplot2::annotate(
    "text",
    x      = threshold,
    y      = Inf,
    label  = round(threshold, 1),
    vjust  = 1.5,
    hjust  = -0.2,
    colour = "steelblue",
    size   = 3.5
  ) +
  ggplot2::geom_text(
    data    = cat_counts,
    mapping = ggplot2::aes(x = x, y = Inf, label = paste0(VulnerabilityClass, "\nn = ", n)),
    vjust   = 3,
    size    = 3.5,
    colour  = "grey30"
  ) +
  ggplot2::theme_minimal()
```

