---
title: "Developing composite indicators for species risk assessment"
subtitle: "A pilot study of avian risk in Gawler Range and Braemar"
author:
- Department for Environment and Water
output:
  bookdown::gitbook:
    split_bib: false
    toc_depth: 3
    fig_caption: true
    reference_docx: "Styles.docx"
link-citations: yes
always_allow_html: yes
csl: "bib_style.csl"
bibliography:
- "all_refs.bib"
---

```{r tars, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE}

library(targets)

# these tars only work inside a knit (otherwise use _targets.yaml directly)
tars <- yaml::read_yaml(here::here("_targets.yaml")) |>
  purrr::modify_tree(leaf = \(x) gsub("\\.\\.\\/\\.\\.", "../../..", x))

```

```{r setup, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE}

# # load targets -------
tar_load_everything(store = tars$report_prep$store)
tar_load_everything(store = tars$envBird$database$store)
tar_load_everything(store = tars$envBird$birdcoin$store)
tar_load_everything(store = tars$birdraster$store)

#--------knitr options--------

knitr::opts_chunk$set(echo = FALSE
                      , warning = FALSE
                      , error = FALSE
                      , message = FALSE
                      , tidy = TRUE
                      , tidy.opts = list(comment = FALSE)
                      , eval.after = "fig.cap" 
                      , output_dir = here::here("report")
                      , progress = FALSE
                      # , dpi = 150
                      , fig.width = 6.5
)

#--------book options-------

options(
  knitr.kable.NA = "",
  knitr.duplicate.label = "allow"
)

# --- tmap options ---

# view/plot mode
#depending on output format
is_html_output <-  function() {
  knitr::opts_knit$get("rmarkdown.pandoc.to")=="html"
}

if(is_html_output()) {
  use_tmap_mode <- "view"
} else {
  use_tmap_mode <- "plot"
  knitr::opts_chunk$set(fig.width = 6.5)
}

##or set manually
if(FALSE){use_tmap_mode <- "plot"}

tmap::tmap_mode(use_tmap_mode)

#suppress progress bars
terra::terraOptions(progress=0)

```

# Aim {-}

The goal of the species risk assessment is to develop a composite indicator of species vulnerability to development pressures, and to translate this into a spatially explicit three-tier zoning map. This map classifies areas as: (1) low risk to biodiversity values and generally suitable for development, (2) moderate risk requiring careful planning and mitigation, and (3) high risk and not suitable for development.

By integrating intrinsic species sensitivity with external pressures (including current, past, and projected impacts), the assessment identifies species most at risk and quantifies the level of pressure they face. This enables the evaluation of both existing and future development impacts.

The resulting spatial outputs are designed to support timely, evidence-based decision-making, guide development planning, and facilitate stakeholder discussions. The framework is flexible and can be adapted to specific development contexts, allowing risks to be reassessed as mitigation measures are incorporated.



