
#### The hierarchy {-}

```{r hierarchy, warning=FALSE, fig.height=8, fig.width=13, out.width="100%"}

plot_coin_hierarchy(h_df)

```

#### Data processing log {-}

- n_minmax: scale using the min and max value of the variable for normalisation, bounded by 0 to 100 (or 1 to 100 for geometric mean) or 0 to 1.

- a_amean: the weighted arithmetic mean for aggregation

- a_gmean: the weighted geometric mean for aggregation

```{r log, echo=FALSE, out.width="90%"}

extract_log_table(coin) %>% 
  kableExtra::kbl(
    escape = FALSE,
    caption = "Extracted data-processing specifications from the coin log. The table summarises the parameters applied at each processing stage and enables comparison across different coins."
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
  ) %>% 
  kableExtra::column_spec(1:4, width = "10em")

```

#### Deconstructing each composite indicator {-}

**Vulnerability**

```{r CI4, echo=FALSE, warning=FALSE, fig.height=5, fig.width=10, out.width="100%"}
plot_coin_bar(coin, "Vulnerability", text.size = 6)
```

**The Sub-Indices**

<style type="text/css">
    div.plotly { height: auto !important;}
</style>

```{r CI3, echo=FALSE, message=FALSE, warning=FALSE, fig.height=15, fig.width=10, out.width="100%"}

## Static plots ------

subindex <- c("Sensitivity", "Exposure", "Pressure")
ncol <- c(3, 4, 3)

plots_static <- Map(function(x, nc) {
  
  p1 <- plot_coin_bar(coin, x)
  p2 <- plot_coin_bar(coin, x, mode = "children", ncol = nc)
  
  patchwork::wrap_plots(p1 / p2) +
    patchwork::plot_layout(heights = c(1, 2)) +
    patchwork::plot_annotation(title = x)
  
}, subindex, ncol)

## Interactive plots ------

plots_interactive <- Map(function(x, nc) {
  
  p1 <- plot_coin_bar(coin, x, interactive = TRUE)
  
  p2 <- plot_coin_bar(
    coin,
    x,
    mode = "children",
    ncol = 3,
    interactive = TRUE
  )
  
  plotly::subplot(
    p1, p2,
    nrows = 2,
    margin = 0.1, # spacing between p1 and p2
    heights = c(0.6, 0.4)
  ) %>%
    plotly::layout(title = list(text = x),
                   height = 700,
                   width = 700)
  
  
}, subindex, ncol)

htmltools::tagList(plots_interactive)
  
```

#### Vulnerability scores {-}

```{r scoretable}

coin$Meta$Unit %>% 
  dplyr::left_join(coin$Data$Aggregated, by = "uCode") %>% 
  dplyr::select(uName, Vulnerability, all_of(subindex)) %>% 
  dplyr::mutate(across(where(is.numeric), round, digits = 2)) %>%
  dplyr::arrange(desc(Vulnerability)) %>% 
    kableExtra::kbl(
    escape = FALSE,
    caption = "Aggregated vulnerability scores and contributing subindices by species, sorted by highest to lowest estimated vulnerability."
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
  ) %>% 
  kableExtra::scroll_box(width = "100%", height = "500px")
  
```

